---
title: "Data Visualization with ggplot2"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prerequisites
Load the tidyverse.

```{r}
library(tidyverse)
```

## Load mpg data
I am putting it into a df called mpg and using ?mpg to get an idea of the structure.

```{r}
mpg<-mpg
?mpg
str(mpg)
mpg
```

## First Basic Plots

Displacement vs hwy mpg
```{r, fig.width=6, fig.height=4}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

Hwy mpg vs displacement
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = cyl))
```

Class vs drv
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
```

Displacement vs hwy mpg colored by class.
Note the change in the legend.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color=class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color=class=="compact"))+
  #theme(legend.position = "none")
  labs(color="Type")+
  scale_color_manual(labels=c("Not Compact", "Compact"), values= c("red", "blue"))
```

Displacement vs hwy mpg scaled by size, alpha, shape, and color
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size=class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha=class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape=class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color="blue")
```

## Faceting

Example facet wrap plots (discrete variable).
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow=2)
```

Example facet grid plots (two variables).
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv~cyl)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(.~cyl)
```

Faceting questions.
Good reference is http://ggplot2.tidyverse.org/
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv~.)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(.~cyl)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(.~class)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow=2, scales="free")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ cyl+drv, nrow=2)
```