---
title: "Homework 3"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Load the Libraries
####nycflights13

```{r, include=FALSE}
library(nycflights13)
library(tidyverse)
```

####Have a look at the data

```{r}
?flights
flights
```

####Filter is by rows
All flights on Jan 1.
```{r}
filter(flights, month==1, day==1)
jan1<-filter(flights, month==1, day==1)
```

All flights on Dec 25, to a variable dec25.
```{r}
(dec25<-filter(flights, month==12, day==25))
```

####Logical operators
```{r}
nov_dec1<-filter(flights, month==11 | month==12)
nov_dec2<-filter(flights, month %in% c(11,12))
```

###Exercises
1. Find all flights that:

a. Had an arrival delay of two or more hours

```{r}
two_hour_delay1<-filter(flights, (arr_delay >120 | dep_delay > 120))
two_hour_delay2<-filter(flights, arr_delay <= 120, dep_delay <= 120)
```

b. Flew to Houston (IAH or HOU)

```{r}
two_hour_delay3<-filter(flights, arr_delay > 120)
str(flights)
houston1<-filter(flights, dest =="IAH" | dest == "HOU")
houston2<-filter(flights, dest %in% c("IAH", "HOU"))
```

c. Were operated by United, American, or Delta

```{r}
str(flights)
carriers1<-filter(flights, carrier == "AA" | carrier =="DL" | carrier == "UA")
carriers2<-filter(flights, carrier %in% c("AA", "DL", "UA"))
```

d. Departed in summer (July, August, and September)

```{r}
filter(flights, between(month, 7, 9))
```

e. Arrived more than two hours late, but didn’t leave late

```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```

f. Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
filter(flights, dep_delay-arr_delay > 30, dep_delay >= 60)
```

g. Departed between midnight and 6am (inclusive)

```{r}
filter(flights, dep_time <=600 | dep_time == 2400)
```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

It picks between a range of values.

```{r}
filter(flights, between(month, 7, 9))
summary(flights)
filter(flights, between(dep_time, 0, 600))
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
summary(flights)
filter(flights, is.na(dep_time))
```

I think these are canceled flights.

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

Need help on this one.

####Arrange rows with arrange()
```{r}
arrange(flights, year, month, day)
arrange(flights, desc(arr_delay))
```

####Exercises

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
arrange(flights, desc(is.na(dep_time)), dep_time)
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, desc(dep_delay))
```

3. Sort flights to find the fastest flights.
```{r}
arrange(flights, air_time)
```

4. Which flights travelled the longest? Which travelled the shortest?
```{r}
arrange(flights, desc(distance))
arrange(flights, distance)
```

####Select is for columns
```{r}
select(flights, year, month, day)
```

```{r}
select(flights, year:day)
```

```{r}
select(flights, -(year:day))
```

```{r}
rename(flights, tail_num= tailnum)
select(flights, time_hour, air_time, everything())
```

####Exercises

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
flights
select(flights, year, month, dep_time, dep_delay, distance)
select(flights, origin, dest, air_time, distance)
```

2. What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, year, month, day, year, year)
```

3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```

####Add new variables with mutate

```{r}
flights_sm<-select(flights, year:day, ends_with("delay"), distance, air_time)
mutate(flights_sm, gain=arr_delay - dep_delay, speed = distance/ air_time*60)
```

####Exercises

1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
flights<-flights
mutate(flights, dep_time2=dep_time/100*60)
mutate(flights, sched_dep_time2=sched_dep_time/100*60)
```

2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

Are these international flights?

```{r}
mutate(flights, test=arr_time-dep_time)
```

3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

departure time-scheduled departure time= delay?

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r}
rank<-mutate(flights, delay.rank = min(desc(dep_delay)))
rank2<-arrange(rank, delay.rank)
filter(rank2, delay.rank<=10)
```

5. What does 1:3 + 1:10 return? Why?

```{r}




```

6. What trigonometric functions does R provide?

```{r}




```