---
title: "Homework 3"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Load the Libraries
####nycflights13

```{r, include=FALSE}
library(nycflights13)
library(tidyverse)
```

####Have a look at the data

```{r}
?flights
flights<-flights
flights
```

####Filter is by rows
All flights on Jan 1.
```{r}
filter(flights, month==1, day==1)
jan1<-filter(flights, month==1, day==1)
```

All flights on Dec 25, to a variable dec25.
```{r}
(dec25<-filter(flights, month==12, day==25))
```

####Logical operators
```{r}
nov_dec1<-filter(flights, month==11 | month==12)
nov_dec2<-filter(flights, month %in% c(11,12))
```

###Exercises
1. Find all flights that:

a. Had an arrival delay of two or more hours

```{r}
two_hour_delay1<-filter(flights, (arr_delay >120 | dep_delay > 120))
two_hour_delay2<-filter(flights, arr_delay <= 120, dep_delay <= 120)
```

b. Flew to Houston (IAH or HOU)

```{r}
two_hour_delay3<-filter(flights, arr_delay > 120)
str(flights)
houston1<-filter(flights, dest =="IAH" | dest == "HOU")
houston2<-filter(flights, dest %in% c("IAH", "HOU"))
```

c. Were operated by United, American, or Delta

```{r}
str(flights)
carriers1<-filter(flights, carrier == "AA" | carrier =="DL" | carrier == "UA")
carriers2<-filter(flights, carrier %in% c("AA", "DL", "UA"))
```

d. Departed in summer (July, August, and September)

```{r}
filter(flights, between(month, 7, 9))
```

e. Arrived more than two hours late, but didn’t leave late

```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```

f. Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
filter(flights, dep_delay-arr_delay > 30, dep_delay >= 60)
```

g. Departed between midnight and 6am (inclusive)

```{r}
filter(flights, dep_time <=600 | dep_time == 2400)
```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

It picks between a range of values.

```{r}
filter(flights, between(month, 7, 9))
summary(flights)
filter(flights, between(dep_time, 0, 600))
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
summary(flights)
filter(flights, is.na(dep_time))
```

I think these are canceled flights.

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

Need help on this one.

####Arrange rows with arrange()
```{r}
arrange(flights, year, month, day)
arrange(flights, desc(arr_delay))
```

####Exercises

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
arrange(flights, desc(is.na(dep_time)), dep_time)
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, desc(dep_delay))
arrange(flights, dep_time)
```

3. Sort flights to find the fastest flights.
```{r}
arrange(flights, air_time)
```

4. Which flights travelled the longest? Which travelled the shortest?
```{r}
arrange(flights, desc(distance))
arrange(flights, distance)
```

####Select is for columns
```{r}
select(flights, year, month, day)
```

```{r}
select(flights, year:day)
```

```{r}
select(flights, -(year:day))
```

```{r}
rename(flights, tail_num= tailnum)
select(flights, time_hour, air_time, everything())
```

####Exercises

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
flights
select(flights, year, month, dep_time, dep_delay, distance)
select(flights, origin, dest, air_time, distance)
```

2. What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, year, month, day, year, year)
```

3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```

####Add new variables with mutate

```{r}
flights_sm<-select(flights, year:day, ends_with("delay"), distance, air_time)
mutate(flights_sm, gain=arr_delay - dep_delay, speed = distance/ air_time*60)
```

####Exercises

1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
flights<-flights
mutate(flights, dep_time2=dep_time/100*60)
mutate(flights, sched_dep_time2=sched_dep_time/100*60)
```

2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

Are these international flights?

```{r}
mutate(flights, test=arr_time-dep_time)
```

3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

departure time-scheduled departure time= delay?

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r}
rank<-mutate(flights, delay.rank = min(desc(dep_delay)))
rank2<-arrange(rank, delay.rank)
filter(rank2, delay.rank<=10)
```

5. What does 1:3 + 1:10 return? Why?

There is a different number of columns.

```{r}
1:3 + 1:10
```

6. What trigonometric functions does R provide?
sin, cos, tan , etc.

####Grouped summaries with summarise()

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

summarise() is not terribly useful unless we pair it with group_by(). This changes the unit of analysis from the complete dataset to individual groups. Then, when you use the dplyr verbs on a grouped data frame they’ll be automatically applied “by group”. For example, if we applied exactly the same code to a data frame grouped by date, we get the average delay per date:

```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

####5.6.1 Combining multiple operations with the pipe
Imagine that we want to explore the relationship between the distance and average delay for each location. Using what you know about dplyr, you might write code like this:

```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")

ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

There’s another way to tackle the same problem with the pipe, %>%:

...As suggested by this reading, a good way to pronounce %>% when reading code is “then”.
```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

####5.6.2 Missing values
You may have wondered about the na.rm argument we used above. What happens if we don’t set it?

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

We get a lot of missing values! That’s because aggregation functions obey the usual rule of missing values: if there’s any missing value in the input, the output will be a missing value. Fortunately, all aggregation functions have an na.rm argument which removes the missing values prior to computation:

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```

In this case, where missing values represent cancelled flights, we could also tackle the problem by first removing the cancelled flights. We’ll save this dataset so we can reuse in the next few examples.

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

####5.6.3 Counts

Whenever you do any aggregation, it’s always a good idea to include either a count (n()), or a count of non-missing values (sum(!is.na(x))). That way you can check that you’re not drawing conclusions based on very small amounts of data. For example, let’s look at the planes (identified by their tail number) that have the highest average delays:

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )
delays

ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```

Wow, there are some planes that have an average delay of 5 hours (300 minutes)!

The story is actually a little more nuanced. We can get more insight if we draw a scatterplot of number of flights vs. average delay:

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```

Not surprisingly, there is much greater variation in the average delay when there are few flights. The shape of this plot is very characteristic: whenever you plot a mean (or other summary) vs. group size, you’ll see that the variation decreases as the sample size increases.

When looking at this sort of plot, it’s often useful to filter out the groups with the smallest numbers of observations, so you can see more of the pattern and less of the extreme variation in the smallest groups. This is what the following code does, as well as showing you a handy pattern for integrating ggplot2 into dplyr flows. It’s a bit painful that you have to switch from %>% to +, but once you get the hang of it, it’s quite convenient.

```{r}

delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

delays

delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```

There’s another common variation of this type of pattern. Let’s look at how the average performance of batters in baseball is related to the number of times they’re at bat. Here I use data from the Lahman package to compute the batting average (number of hits / number of attempts) of every major league baseball player.

When I plot the skill of the batter (measured by the batting average, ba) against the number of opportunities to hit the ball (measured by at bat, ab), you see two patterns:

As above, the variation in our aggregate decreases as we get more data points.

There’s a positive correlation between skill (ba) and opportunities to hit the ball (ab). This is because teams control who gets to play, and obviously they’ll pick their best players.

```{r}
library(Lahman)

batting <- as_tibble(Lahman::Batting)

batters <- batting %>% 
  group_by(playerID) %>% 
  summarise(
    ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    ab = sum(AB, na.rm = TRUE)
  )

batters

batters %>% 
  filter(ab > 100) %>% 
  ggplot(mapping = aes(x = ab, y = ba)) +
    geom_point() + 
    geom_smooth(se = FALSE)
```

This also has important implications for ranking. If you naively sort on desc(ba), the people with the best batting averages are clearly lucky, not skilled:
```{r}
batters %>% 
  arrange(desc(ba))
```

####5.6.4 Useful summary functions

Just using means, counts, and sum can get you a long way, but R provides many other useful summary functions:

Measures of location: we’ve used mean(x), but median(x) is also useful. The mean is the sum divided by the length; the median is a value where 50% of x is above it, and 50% is below it.

It’s sometimes useful to combine aggregation with logical subsetting. We haven’t talked about this sort of subsetting yet, but you’ll learn more about it in subsetting.

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0]) # the average positive delay
  )
```

Measures of spread: sd(x), IQR(x), mad(x). The mean squared deviation, or standard deviation or sd for short, is the standard measure of spread. The interquartile range IQR() and median absolute deviation mad(x) are robust equivalents that may be more useful if you have outliers.

```{r}
# Why is distance to some destinations more variable than to others?
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```

Measures of rank: min(x), quantile(x, 0.25), max(x). Quantiles are a generalisation of the median. For example, quantile(x, 0.25) will find a value of x that is greater than 25% of the values, and less than the remaining 75%.

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )
```

Measures of position: first(x), nth(x, 2), last(x). These work similarly to x[1], x[2], and x[length(x)] but let you set a default value if that position does not exist (i.e. you’re trying to get the 3rd element from a group that only has two elements). For example, we can find the first and last departure for each day:

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time), 
    last_dep = last(dep_time)
  )
```

These functions are complementary to filtering on ranks. Filtering gives you all variables, with each observation in a separate row:

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

Counts: You’ve seen n(), which takes no arguments, and returns the size of the current group. To count the number of non-missing values, use sum(!is.na(x)). To count the number of distinct (unique) values, use n_distinct(x).

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```

Counts are so useful that dplyr provides a simple helper if all you want is a count:

```{r}
not_cancelled %>% 
  count(dest)
```

You can optionally provide a weight variable. For example, you could use this to “count” (sum) the total number of miles a plane flew:

```{r}
not_cancelled %>% 
  count(tailnum, wt = distance)
```

Counts and proportions of logical values: sum(x > 10), mean(y == 0). When used with numeric functions, TRUE is converted to 1 and FALSE to 0. This makes sum() and mean() very useful: sum(x) gives the number of TRUEs in x, and mean(x) gives the proportion.

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))

# What proportion of flights are delayed by more than an hour?
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_perc = mean(arr_delay > 60))
```

####5.6.5 Grouping by multiple variables

When you group by multiple variables, each summary peels off one level of the grouping. That makes it easy to progressively roll up a dataset:

```{r}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))

(per_month <- summarise(per_day, flights = sum(flights)))

(per_year  <- summarise(per_month, flights = sum(flights)))
```

Be careful when progressively rolling up summaries: it’s OK for sums and counts, but you need to think about weighting means and variances, and it’s not possible to do it exactly for rank-based statistics like the median. In other words, the sum of groupwise sums is the overall sum, but the median of groupwise medians is not the overall median.

####5.6.6 Ungrouping

If you need to remove grouping, and return to operations on ungrouped data, use ungroup().
```{r}
daily %>% 
  ungroup() %>%             # no longer grouped by date
  summarise(flights = n())  # all flights
```

####5.6.7 Exercises

1. Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.

A flight is always 10 minutes late.

A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.

99% of the time a flight is on time. 1% of the time it’s 2 hours late.

Which is more important: arrival delay or departure delay?

arrival delay because you can make up time en route.

```{r}
by_month <- group_by(flights, month)
summarise(by_month, delay = mean(dep_delay, na.rm = TRUE))
summarise(by_month, delay = mean(arr_delay, na.rm = TRUE))
```

2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).

```{r}
not_cancelled %>%
  group_by(tailnum) %>%
  summarize(flights= sum(distance))
           
not_cancelled %>%
  group_by (dest) %>%
  summarize(flights= sum(distance))
```

3. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?

Can you arrive if you don't depart? I suppose you could depart and then not arrive?!

```{r}
filter(flights, !is.na(dep_delay), is.na(arr_delay)) %>%
  select(dep_time, arr_time, sched_arr_time, dep_delay, arr_delay, air_time)
```

4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}
cancelled<- flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(per_cancelled = mean(cancelled) * 100,
            ave_dep_delay = mean(dep_delay, na.rm = TRUE))
cancelled

ggplot(data=cancelled, aes(x = ave_dep_delay, y=per_cancelled))  +
  geom_point() +
  geom_smooth()
```

5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))
```{r}
flights %>%
  group_by(carrier) %>%
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(arr_delay))
```

6. What does the sort argument to count() do. When might you use it?



####5.7 Grouped mutates (and filters)

Grouping is most useful in conjunction with summarise(), but you can also do convenient operations with mutate() and filter():

Find the worst members of each group:

```{r}
flights_sml <- flights %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
```

Find all groups bigger than a threshold:

```{r}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```

Standardise to compute per group metrics:

```{r}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```

A grouped filter is a grouped mutate followed by an ungrouped filter. I generally avoid them except for quick and dirty manipulations: otherwise it’s hard to check that you’ve done the manipulation correctly.

Functions that work most naturally in grouped mutates and filters are known as window functions (vs. the summary functions used for summaries). You can learn more about useful window functions in the corresponding vignette: vignette("window-functions").

####5.7.1 Exercises

1. Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.

2. Which plane (tailnum) has the worst on-time record?

```{r}
?arrange
worst<-flights %>%
  group_by(tailnum) %>% 
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE))
  arrange(worst, desc(arr_delay))

worst2<-flights %>%
  group_by(tailnum) %>% 
  filter(arr_delay>0) %>%
  count(tailnum) %>%
  arrange(desc(n))
worst2
```

3. What time of day should you fly if you want to avoid delays as much as possible?

4. For each destination, compute the total minutes of delay. For each, flight, compute the proportion of the total delay for its destination.

5. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag() explore how the delay of a flight is related to the delay of the immediately preceding flight.

```{r}
not_cancelled %>%
  mutate(
    previous_dep_delay = dep_delay,
    next_dep_delay = lag(previous_dep_delay)
  ) %>%
  ggplot(mapping = aes(x = previous_dep_delay, y = next_dep_delay)) + 
  geom_point() +
  geom_smooth()
```

6. Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.

8. For each plane, count the number of flights before the first delay of greater than 1 hour.
