---
title: "ggplot (con’t)"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load the libraries
```{r}
library(tidyverse)
library(maps)
```

##3.6 Geometric Objects

```{r}
# scatterplot
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# fitted line
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

This is pretty helpful.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```

Thinking about geom's in layers.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```

Filter is a dplyr command.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
line; depends on the data? geom_abline()
boxplot; geom_boxplot()
area; geom_area()

2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

3. What does show.legend = FALSE do? What happens if you remove it?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv, show.legend = FALSE)) + 
  geom_point() + 
  geom_smooth()
```

4. What does the se argument to geom_smooth() do?
Confidence intervals go away? There is too much overlap.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth()
```

5. Will these two graphs look different? Why/why not?
No, the first one universally defines the variables.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

6. Recreate the R code necessary to generate the following graphs.

Plot 1
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

Plot 2
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

Plot 3
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

Plot 4
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

Plot 5
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, color = drv, linetype = drv), se= FALSE)
```

Plot 6
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "white", size = 3)
  geom_point(mapping = aes(x = displ, y = hwy, color = drv), size = 3)
```

##3.7 Statistical Transformations

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

Same plot, specifying stat_count
```{r}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```

```{r}
demo <- tribble(
  ~a,      ~b,
  "bar_1", 20,
  "bar_2", 30,
  "bar_3", 40
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = a, y = b), stat = "identity")
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

```{r}
?stat_summary
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

geom_pointrange

```{r}
?diamonds
?stat_bin
?stat_summary
?geom_pointrange
geom_pointrange()
ggplot(data = diamonds) + 
  geom_pointrange(mapping = aes(x = cut, y = depth), stat = "summary", 
  fun.ymin = min,
  fun.ymax = max,
  fun.y = median)
```

2. What does geom_col() do? How is it different to geom_bar()?

geom_bar uses the count transformation.
geom_col uses identity transformation.

```{r}
?stat_identity
diamonds

ggplot(data = diamonds) + 
  geom_col(mapping = aes(x = cut, y = price))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = price))
```
```

3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?


4. What variables does stat_smooth() compute? What parameters control its behaviour?

same as geom_smooth

```{r}
?stat_smooth
```

5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

All of the bars are of the same height. 
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
  
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop.., group = color))
```

##3.8 Position Adjustments

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```
 
```{r} 
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
  
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

Identity
```{r} 
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = .4, position = "identity")

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1, position = "dodge")

ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")
```

Fill
```{r} 
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

Dodge
```{r} 
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

Jitter to avoid overplotting
```{r} 
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

1. What is the problem with this plot? How could you improve it?
Overplotting, use jitter.

```{r} 
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
  
 ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter() 
```

2. What parameters to geom_jitter() control the amount of jittering?

height and width
size

```{r} 
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = position_jitter(width = 1, height = 2))
  
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = position_jitter(width = 1, height = 2))
  
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter(size=1)
```

3. Compare and contrast geom_jitter() with geom_count().

```{r} 
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()
  
 ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count(alpha=0.3) 
```

4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

Dodge

```{r}
mpg
ggplot(data = mpg, aes(x = class, y = hwy, fill = trans)) +
  geom_boxplot()
  
ggplot(data = mpg, aes(x = drv, y = hwy, fill = manufacturer)) +
  geom_boxplot(position=position_dodge(1))
```

##3.9 Coordinate Systems

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

```{r}
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```

1. Turn a stacked bar chart into a pie chart using coord_polar().

```{r}
diamonds
ggplot(diamonds, aes(x = cut, fill = color)) +
  geom_bar()

ggplot(data=diamonds)
geom_bar(mapping=aes(x=cut, fill=clarity), position = "fill")
coord_polar()

ggplot(diamonds, aes(x = cut, fill = color)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y")
```

2. What does labs() do? Read the documentation.

Adds labels.

3. What’s the difference between coord_quickmap() and coord_map()?


4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
  
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline()
```
